---
title: "Ex2 App Marketing Structure"
output: 
  html_notebook:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(ggplot2)

# Clear All Variables & Clear the Screen
rm(list=ls())
cat("\014")

# Read in the Data
df = read.csv("Ex2_Data_R.csv")

# Explore the data
#str(df)
#summary(df)
```
## Prepare Data
```{r}
# Convert unit sales to log unit sales
df$logUNITS1 <- log(df$UNITS1)
df$logUNITS2 <- log(df$UNITS2)
df$logUNITS3 <- log(df$UNITS3)

```


# Using a linear regressions framework with unit sales (and/or log unit sales) as the dependent variable, investigate the effect of regular price, feature and rating on sales for each of the three apps. How good are these models? 

## Simple Regression Model for App1 (only consider own price, feature and rating) {#app1.simple}
```{r}
# Linear model
lm.linear.smp.1 = lm(UNITS1~REGPR1+FEAT1+RATING1, df)
# semi-Log model
lm.log.smp.1 = lm(logUNITS1~REGPR1+FEAT1+RATING1, df)

# Calculate price elasticity for the linear model
price.coeff.linear.smp.1 = unname(lm.linear.smp.1$coefficients[2], force = FALSE)
elas.linear.smp.1        = price.coeff.linear.smp.1*mean(df$REGPR1)/mean(df$UNITS1)

# Calculate price elasticity for the Semi-log model
price.coeff.log.smp.1 = unname(lm.log.smp.1$coefficients[2], force = FALSE)
elas.log.smp.1        = price.coeff.log.smp.1*mean(df$REGPR1)

summary(lm.linear.smp.1)
summary(lm.log.smp.1)
```
**Comment:**
For App 1, both the simple models fit quite well and have a high R-square despite not including competitive effects. All three variables are significant. 

## Simple Regression Model for App2 (only consider own price, feature and rating) {#app2.simple}
```{r}
# Linear model
lm.linear.smp.2 = lm(UNITS2~REGPR2+FEAT2+RATING2, df)
# Semi-Log model
lm.log.smp.2 = lm(logUNITS2~REGPR2+FEAT2+RATING2, df)

# Calculate price elasticity for the linear model
price.coeff.linear.smp.2 = unname(lm.linear.smp.2$coefficients[2], force = FALSE)
elas.linear.smp.2        = price.coeff.linear.smp.2*mean(df$REGPR2)/mean(df$UNITS2)

# Calculate price elasticity for the Semi-log model
price.coeff.log.smp.2 = unname(lm.log.smp.2$coefficients[2], force = FALSE)
elas.log.smp.2        = price.coeff.log.smp.2*mean(df$REGPR2)

summary(lm.linear.smp.2)
summary(lm.log.smp.2)
```
**Comment:**
The simple linear model fits better with a R-square of 0.67. All three variables are significant.

## Simple Regression Model for App3 (only consider own price, feature and rating) {#app3.simple}
```{r}
# Linear model
lm.linear.smp.3 = lm(UNITS3~REGPR3+FEAT3+RATING3, df)
# Semi-Log model
lm.log.smp.3 = lm(logUNITS3~REGPR3+FEAT3+RATING3, df)

# Calculate price elasticity for the linear model
price.coeff.linear.smp.3 = unname(lm.linear.smp.3$coefficients[2], force = FALSE)
elas.linear.smp.3        = price.coeff.linear.smp.3*mean(df$REGPR3)/mean(df$UNITS3)

# Calculate price elasticity for the Semi-log model
price.coeff.log.smp.3 = unname(lm.log.smp.3$coefficients[2], force = FALSE)
elas.log.smp.3        = price.coeff.log.smp.3*mean(df$REGPR3)

summary(lm.linear.smp.3)
summary(lm.log.smp.3)
```
**Comment:**
Neither model fits particularly well as the R-squares are relatively small. The simple linear model only includes the feature varibale and leaves out the price. The semi-log one accounts for more variables but has a lower R-square.


# An important source of variation in sales often comes from competitive marketing activity. Investigate the impact on sales of each app from the changes in the marketing activity of competing apps. What, if any, competitive terms would you want to include in your final models? 

## Full Regression Model for App1  (consider both own and competitors' price, feature and rating) {#app1.full}
```{r}
# Linear model for app1 
lm.linear.1 = lm(UNITS1~REGPR1+FEAT1+RATING1+REGPR2+FEAT2+RATING2+REGPR3+FEAT3+RATING3, df)
# Semi-Log model for app1
lm.log.1 = lm(logUNITS1~REGPR1+FEAT1+RATING1+REGPR2+FEAT2+RATING2+REGPR3+FEAT3+RATING3, df)

# Calculate price elasticity for linear model
price.coeff.linear.1 = unname(lm.linear.1$coefficients[2], force = FALSE)
elas.linear.1        = price.coeff.linear.1*mean(df$REGPR1)/mean(df$UNITS1)

# Calculate price elasticity for Semi-log model
price.coeff.log.1 = unname(lm.log.1$coefficients[2], force = FALSE)
elas.log.1        = price.coeff.log.1*mean(df$REGPR1)

summary(lm.linear.1)
summary(lm.log.1)
```

## Full Regression Model for App2  (consider both own and competitors' price, feature and rating) {#app2.full}
```{r}
# Linear model for app2
lm.linear.2 = lm(UNITS2~REGPR1+FEAT1+RATING1+REGPR2+FEAT2+RATING2+REGPR3+FEAT3+RATING3, df)
# Semi-Log model for app2
lm.log.2 = lm(logUNITS2~REGPR1+FEAT1+RATING1+REGPR2+FEAT2+RATING2+REGPR3+FEAT3+RATING3, df)

# Calculate price elasticity for linear model
price.coeff.linear.2 = unname(lm.linear.2$coefficients[5], force = FALSE)
elas.linear.2        = price.coeff.linear.2*mean(df$REGPR2)/mean(df$UNITS2)

# Calculate price elasticity for Semi-log model
price.coeff.log.2 = unname(lm.log.2$coefficients[5], force = FALSE)
elas.log.2        = price.coeff.log.2*mean(df$REGPR2)

summary(lm.linear.2)
summary(lm.log.2)
```

## Full Regression Model for App3  (consider both own and competitors' price, feature and rating) {#app3.full}
```{r}
# Linear model for app3
lm.linear.3 = lm(UNITS3~REGPR1+FEAT1+RATING1+REGPR2+FEAT2+RATING2+REGPR3+FEAT3+RATING3, df)
# Semi-Log model for app3
lm.log.3 = lm(logUNITS3~REGPR1+FEAT1+RATING1+REGPR2+FEAT2+RATING2+REGPR3+FEAT3+RATING3, df)

# Calculate price elasticity for linear model
price.coeff.linear.3 = unname(lm.linear.3$coefficients[8], force = FALSE)
elas.linear.3        = price.coeff.linear.3*mean(df$REGPR3)/mean(df$UNITS3)

# Calculate price elasticity for Semi-log model
price.coeff.log.3 = unname(lm.log.3$coefficients[8], force = FALSE)
elas.log.3        = price.coeff.log.3*mean(df$REGPR3)

summary(lm.linear.3)
summary(lm.log.3)
```


# Propose the “best” regression model for each app, taking into account own-effects and competitive-effects. Comment on the quality of these final models.

## App1
```{r}
# Compare models for App1
#summary(lm.linear.smp.1)  
#summary(lm.log.smp.1)
#summary(lm.linear.1)
#summary(lm.log.1) # best model: High R-square, taken into account of competitor's price
```
[own models](#app1.simple),
[own&comp models](#app1.full)  
We propose the 'semi-log own&comp'(lm.log.1) as the best model.
Though the 'linear own&comp' has the highest R-square, it doesn't take competitor's prices into account.
The semi-log has not only app1's price and feature significant, but also app2's price.

## App2  
```{r}
# Compare models for App2
#summary(lm.linear.smp.2) 
#summary(lm.log.smp.2)
#summary(lm.linear.2)  # best model: Highest R-square, incorporates the cross-price effect by competitor(app1) and the effect of app 1's feature. 
#summary(lm.log.2)
```
[own models](#app2.simple), 
[own&comp models](#app2.full)  
We propose the 'linear own&comp'(lm.linear.1) as the best model.
Not only it has the highest R-Square, it has taken App1's price and feature into account.

## App3
```{r}

# Compare models for App3
#summary(lm.linear.smp.3)
#summary(lm.log.smp.3)
#summary(lm.linear.3)
#summary(lm.log.3)    # best model: Highest R-square, incorporate the cross-price effect by app1
```
[own models](#app3.simple), 
[own&comp models](#app3.full)  
We propose the 'semi-log own&comp'(lm.log.3) as the best model.
Not only it has the highest R-Square, it has taken App1's price into account.


# What do these best models tell you about market structure and inter-app competition in this game category, e.g., create a clout/vulnerability map (as we did in class)?

## Construct a price coefficient table
```{r}
# Build a table of price coefficients with cross-effects
# The selected model is named by: model_type.app.app_number
# Also, only significant price coefficients are selected
semi_log.app.1 <- c(lm.log.1$coefficients[2], lm.log.1$coefficients[5], NA)
linear.app.2 <- c(lm.linear.2$coefficients[2], lm.linear.2$coefficients[5], NA)
semi_log.app.3 <- c(lm.log.3$coefficients[2], NA, lm.log.3$coefficients[8])

price_coeff_table <- data.frame(semi_log.app.1,linear.app.2,semi_log.app.3)
row.names(price_coeff_table) <- c("price.coeff.app.1", "price.coeff.app.2", "price.coeff.app.3")
price_coeff_table <- data.frame(t(price_coeff_table))
price_coeff_table
```

## Calculate the cross-price elasticity
```{r}
# Convert price coefficients to elasticity
price_elas_table <- price_coeff_table
colnames(price_elas_table) <-c("price.elas.app.1","price.elas.app.2","price.elas.app.3")

price_elas_table$price.elas.app.1[1] <- price_elas_table$price.elas.app.1[1]*(mean(df$REGPR1))
price_elas_table$price.elas.app.1[2] <- price_elas_table$price.elas.app.1[2]*(mean(df$REGPR1)/mean(df$UNITS2))
price_elas_table$price.elas.app.1[3] <- price_elas_table$price.elas.app.1[3]*mean(df$REGPR1)
price_elas_table$price.elas.app.2[1] <- price_elas_table$price.elas.app.2[1]*(mean(df$REGPR2))
price_elas_table$price.elas.app.2[2] <- price_elas_table$price.elas.app.2[2]*(mean(df$REGPR2)/mean(df$UNITS2))
price_elas_table$price.elas.app.3[3] <- price_elas_table$price.elas.app.3[3]*mean(df$REGPR3)
price_elas_table
```

## Build the clout and vulnerability table
```{r}
# Build a clout and vulnerability table by adding the clout and vulnerability up for each app
clout.app.1 <- price_elas_table$price.elas.app.1[2]+price_elas_table$price.elas.app.1[3]
clout.app.2 <- price_elas_table$price.elas.app.2[1]
clout.app.3 <- 0

vulnerability.app.1 <- price_elas_table$price.elas.app.2[1]
vulnerability.app.2 <- price_elas_table$price.elas.app.1[2]
vulnerability.app.3 <- price_elas_table$price.elas.app.1[3]

clout_and_vulnerability <- data.frame(c("app1","app2", "app3"),
                                      c(clout.app.1,clout.app.2,clout.app.3), 
                                      c(vulnerability.app.1,vulnerability.app.2,vulnerability.app.3))
colnames(clout_and_vulnerability) <- c("app", "clout", "vulnerability")
clout_and_vulnerability
```

## Visualize clout and vulnerability 
```{r}
# add market share column
# market share is calculated by the percentage of average unit size
avg.unit.1 <- mean(df$UNITS1)
avg.unit.2 <- mean(df$UNITS2)
avg.unit.3 <- mean(df$UNITS3)
total.unit <- avg.unit.1+avg.unit.2+avg.unit.3
clout_and_vulnerability$share <- c(100*avg.unit.1/total.unit, 100*avg.unit.2/total.unit, 100*avg.unit.3/total.unit)

# plot the clout & vulnerability chart
ggplot(data=clout_and_vulnerability, mapping=aes(x=vulnerability, y=clout, label=app)) + geom_point(aes(size=share), color="coral1") + geom_text(nudge_x=0.4) + ggtitle("Clout and Vulnerability")
```
**comment on clout and vulnerability**  
App1 has strong clout and little vulnerability. This means a price deduction in App1 will largely steal the share of other apps. App2 and App3 both have little to none clout, meaning the price change of these two apps have little effect on competitors. However, since App3 has a high vulnerability, competitors' price drop will sharply decreases App3's share.  
  
App1 with the highest price and greatest market share is a premium brand. It has larger effects on other two small apps than vice versa. 



